---
title: "tidyverse pt. 1"
author: "Nicholas Chung"
date: "12/08/2019"
output: 
  html_document:
    code_folding: hide
    highlight: pygments
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, message=FALSE}
library(tidyverse)
library(knitr)
```

# Assignment
In this assignment, you’ll practice collaborating around a code project with GitHub.  You could consider our collective work as building out a book of examples on how to use TidyVerse functions.

GitHub repository:  https://github.com/acatlin/FALL2019TIDYVERSE
FiveThirtyEight.com datasets.
Kaggle datasets. 

You have two tasks:

Create an Example.  Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset. (25 points)
Extend an Existing Example.  Using one of your classmate’s examples (as created above), extend his or her example with additional annotated code. (15 points)
You should clone the provided repository.  Once you have code to submit, you should make a pull request on the shared repository.  Minimally, you should be submitted .Rmd files; ideally, you should also submit an .md file and update the README.md file with your example.

# Dataset
For this Tidyverse assignment, I've selected this Kaggle dataset: https://www.kaggle.com/akhilv11/border-crossing-entry-data

```{r load_data}
# load csv
raw.df <- read.csv("Border_Crossing_Entry_Data.csv", header = TRUE, stringsAsFactors = FALSE)
head(raw.df, 10)
```

# Method
## explore data to formulate question
```{r}
# investigate dates
raw.df$Date %>% n_distinct()

# narrow scope of investigation to California
ca <- raw.df %>% 
  filter(State == "California") %>%
  select(Port.Name, Measure, Value, Date)

# visualize california border crossing count
ggplot(ca, aes(x=Date, y=Value)) + 
  geom_point() +
  ggtitle("(n) border crossings, california") +
  xlab("year") + ylab("count")

# group border crossings by port of entry, measure, and border
mx.border <- raw.df %>%
  filter(Border == "US-Mexico Border") %>%
  select(Port.Name, State, Border, Date, Measure, Value)

# what are the different methods of entry?
mx.border %>% 
  distinct(Measure)
```

## use Tidyverse methods to frame approach to question: "How do different modes of port of entry access compare across US-Mexico border?"
Tidyverse methods utlized:
* filter
* summarize
```{r explore question}
# summary of large non-rail vehicle-based entry
big.wheeled <- c("Trucks", "Buses", "Truck Containers Empty", "Bus Passengers", "Truck Containers Full")
mx.border %>%
  filter(Measure %in% big.wheeled) %>%
  select(Measure, Value) %>%
  summarize(mean = mean(Value))

# summary of small vehicle-based entry
small.wheeled <- c("Personal Vehicles", "Personal Vehicle Passengers")
mx.border %>%
  filter(Measure %in% small.wheeled) %>%
  select(Measure, Value) %>%
  summarize(mean = mean(Value))

# summary of train-based entry
rail.wheeled <- c("Train Passengers", "Trains", "Rail Containers Full", "Rail Containers Empty")
mx.border %>%
  filter(Measure %in% rail.wheeled) %>%
  select(Measure, Value) %>%
  summarize(mean = mean(Value))

# summary of pedestrain entry
no.wheeled <- c("Pedestrians")
mx.border %>%
  filter(Measure %in% no.wheeled) %>%
  select(Measure, Value) %>%
  summarize(mean = mean(Value))
```

# Finding
The above findings suggest that personal vehicles are the most prevalent mode of transportation at ports of entry along the US-Mexico border.
