---
title: "tidyverse pt. 1"
author: "Nicholas Chung"
date: "12/08/2019"
output: 
  html_document:
    code_folding: hide
    highlight: pygments
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, message=FALSE}
library(tidyverse)
library(knitr)
```

# Assignment
In this assignment, you’ll practice collaborating around a code project with GitHub.  You could consider our collective work as building out a book of examples on how to use TidyVerse functions.

GitHub repository:  https://github.com/acatlin/FALL2019TIDYVERSE
FiveThirtyEight.com datasets.
Kaggle datasets. 

You have two tasks:

Create an Example.  Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset. (25 points)
Extend an Existing Example.  Using one of your classmate’s examples (as created above), extend his or her example with additional annotated code. (15 points)
You should clone the provided repository.  Once you have code to submit, you should make a pull request on the shared repository.  Minimally, you should be submitted .Rmd files; ideally, you should also submit an .md file and update the README.md file with your example.

# Extend Existing Example
For this second component, I've chosen to extend ["trump_approval_ratings"](https://github.com/pnojai/FALL2019TIDYVERSE/blob/master/trump_approval_ratings.Rmd) by Donny Lofland

# Method
In the identified example, the methods focus on the demonstration of the dplyr:filter() and dplyr:gather() functions to show the difference in approval and disapproval amongst the subgroup "voters".

Here, we will extend the user of dplyr:filter() to show the mean approval and disapproval for each of the various subgroups.

### load data
```{r}
# function to load data
load_cached_url <- function(filename, url) {
  dir.create(file.path('.', 'data'), showWarnings = FALSE)
  fqfn <- paste('./data', filename, sep = '/')

  # Download the file and save it, if we don't have it yet.  
  if(!file.exists(fqfn)) {
    data <- read_csv(url)
    write_csv(data, fqfn)
  }

  # Guess column specification.
  fqfn_spec <- spec_csv(fqfn)
  
  # Identify date columns.
  # Assumption: Column name ends with "date".
  date_cols <- names(fqfn_spec$cols)[str_detect(names(fqfn_spec$cols),
                                     pattern = "date$")]
  
  # Remove one of the date column names, in order to illustrate later the
  # alternative of specifying datatype after reading, too.
  date_cols <- date_cols[-1]

  # Set specification explicitly for date columns.
  for (column in date_cols) {
    fqfn_spec$cols[[column]] <- col_date("%m/%d/%Y")  
  }
  
  # Read file, setting datatypes for all columns.
  data <- read_csv(fqfn, col_types = fqfn_spec)
  
  return(as_tibble(data))
}
```

```{r read_csv1, eval=TRUE, message=FALSE}

polls <- load_cached_url(
  'trump-approval-ratings-polls.csv', 
  'https://projects.fivethirtyeight.com/trump-approval-data/approval_polllist.csv'
  )

results <- load_cached_url(
  'trump-approval-ratings-results.csv',
  'https://projects.fivethirtyeight.com/trump-approval-data/approval_topline.csv')

head(polls)
head(results)
```

```{r}
# confirm discrete subgroups
results %>%
  distinct(subgroup)

# mean approval per subgroup
results %>% 
  filter(subgroup == "Adults") %>%
  select(subgroup, approve_estimate) %>%
  summarize(m.approve = mean(approve_estimate))

results %>% 
  filter(subgroup == "All polls") %>%
  select(subgroup, approve_estimate) %>%
  summarize(m.approve= mean(approve_estimate))

results %>% 
  filter(subgroup == "Voters") %>%
  select(subgroup, approve_estimate) %>%
  summarize(m.approve = mean(approve_estimate))

# mean disapproval per subgroup
results %>% 
  filter(subgroup == "Adults") %>%
  select(subgroup, disapprove_estimate) %>%
  summarize(m.disapprove = mean(disapprove_estimate))

results %>% 
  filter(subgroup == "All polls") %>%
  select(subgroup, disapprove_estimate) %>%
  summarize(m.disapprove = mean(disapprove_estimate))

results %>% 
  filter(subgroup == "Voters") %>%
  select(subgroup, disapprove_estimate) %>%
  summarize(m.disapprove = mean(disapprove_estimate))
```
